
```{r}

setwd("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data")


# Step 1: load the files 

#install.packages(c("tidyverse", "readxl"))

library(readxl) 
library(dplyr)


Parthenon <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Parthenon')%>%
  filter(if_any(everything(), ~ !is.na(.)))
Parthenon<- Parthenon[1:58,]


Kupad_Shivakumar <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Kupad_Shivakumar')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Kupad_Shivakumar<- Kupad_Shivakumar[1:89,]

Bangladesh <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Bangladesh')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Bangladesh<- Bangladesh[1:29,]


Bandara <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '196_Bandara')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Pfrimer <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '195_Pfrimer')%>%
  filter(if_any(everything(), ~ !is.na(.)))


Davy <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '194_Davy')%>%
  filter(if_any(everything(), ~ !is.na(.)))
Davy<- Davy[1:75,]


Giacomo_Fassini <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '192_Giacomo_Fassini')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Goldman <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '191_Goldman')%>%
  filter(if_any(everything(), ~ !is.na(.)))



weaver <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '190_Weaver(not_completed)')%>%
  filter(if_any(everything(), ~ !is.na(.)))



Desneves <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '193_Desneves')%>%
  filter(if_any(everything(), ~ !is.na(.)))


McGrosky  <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '189_McGrosky')%>%
  filter(if_any(everything(), ~ !is.na(.)))


Holsgrove_West<- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '188_Holsgrove-West')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Holsgrove_West<- Holsgrove_West[1:25,]

silver<- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '187_Silver')%>%
  filter(if_any(everything(), ~ !is.na(.)))

silver<- silver[1:118,]

Pongcharoen <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '186_Pongcharoen')%>%
  filter(if_any(everything(), ~ !is.na(.)))
Pongcharoen <- Pongcharoen[1:497,]


coker_RH <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet= 'TEE+FILE+COKER+RH')%>%
  filter(if_any(everything(), ~ !is.na(.)))


# Step 2: merge all the lists together 

df_lists <- list(
  Bangladesh = Bangladesh,
  Bandara = Bandara,
  Davy = Davy,
  Desneves = Desneves,
  Giacomo_Fassini = Giacomo_Fassini,
  Goldman = Goldman,
  Holsgrove_West = Holsgrove_West,
  Kupad_Shivakumar = Kupad_Shivakumar,
  McGrosky = McGrosky,
  Parthenon = Parthenon,
  Pfrimer = Pfrimer,
  Pongcharoen = Pongcharoen,
  silver = silver,
  weaver = weaver,
  coker_RH = coker_RH)


```

# Q1/ Are all the mandatory variable filled in ? 

```{r}

#Question 1 ---- 

# Q1/ Are all the mandatory variable filled in ? 

req_var <- c(
  "Study ID",
  "Individual ID",
  "Age",
  "Sex",
  "Dosing body weight",
  "ko",
  "kd",
  "No",
  "Nd",
  "rCO2",
  "rCO2 method",
  "TEE")


  for (df_name in names(df_lists)) {                      #Loop over the 14 dataset
    
    current_df <- df_lists[[df_name]]                     
    cat("\n📋 Checking:", df_name, "\n")
    
    for (var in req_var) {                               #2- loop over the req_var and check if the selected data has columnm in req_var
      
      if (var %in% colnames(current_df)) {                    #3- if req_var in one of the df 
        
        values <- current_df[[var]]           # ✅ keep all values including NAs
        
        if (all(!is.na(values))) {            # ✅ check if ALL values are filled
          cat("  \u2705 Fully filled", var, "- Fully filled \n")
        } else {
          n_missing <- sum(is.na(values))     # ✅ count the NAs
          cat("  ⚠️ ", var, "-", n_missing, "missing values \n")
        }
        
      }  else {
        cat("  ❌", var, "- Column not found \n")
      }
    }
  }

```{r}
library(tidyverse)

# Build a results dataframe from your loop
results <- map_dfr(names(df_lists), function(df_name) {
  current_df <- df_lists[[df_name]]
  map_dfr(req_var, function(var) {
    if (!var %in% colnames(current_df)) {                    #If the variable is missing, put absent in status
      tibble(dataset = df_name, variable = var, status = "absent", pct_complete = NA)
    } else {
      values <- current_df[[var]]                      #If not missing take the values
      n_total   <- length(values)                      # count the values 
      n_missing <- sum(is.na(values))                   #Count the missing values 
      tibble(                                            #create results dataframe 
        dataset     = df_name,
        variable    = var,
        status      = ifelse(n_missing == 0, "ok", "warn"),            #If the missing zero put ok, else warn 
        pct_complete = round((1 - n_missing / n_total) * 100)           #count the %
      )
    }
  })
})

# Plot the heatmap
results %>%                              #Take the df results and make label column to 
  mutate(label = case_when(
    status == "absent" ~ "N/A",
    pct_complete == 100 ~ "100%",
    TRUE ~ paste0(pct_complete, "%")     #for all cases write the value pct_com and put %
  )) %>%
  ggplot(aes(x = variable, y = dataset, fill = pct_complete)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "#E24B4A", mid = "#FAC775", high = "#1D9E75",
                       midpoint = 50, na.value = "#D3D1C7", limits = c(0, 100),
                       name = "% complete") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = NULL, title = "QC completeness across datasets")
```
```{r}
# Q2/ Are all autocalculated columns empty? 

empty_col<- c("Database measurement ID","Database individual ID","BMI","Average weight",
              "Weight change","ko/kd","Nd/No","pred DSR","autocompleted CO2 production","TBW","Autocompleted TEE Speakman et al 2021","Water turnover","Hydration",
              "FFM (kg)","FFM(0.732)","FM (kg)","Adiposity","Analysis lab anon")


for( df_name in names(df_lists)){
  df<- df_lists[[df_name]]
  matched_cols<- intersect(colnames(df),empty_col) #find the empty_cols in the datasets
  
  if(length(matched_cols) == 0) next             #skip if the matched _cols == zero 
  cat('\n',df_name,'\n')
   cat('---------------------','\n')
  for(col in matched_cols){
    values<-df[[col]]                           #extract the column values
    
    if(all(is.na(values) | values == "")) {
      
      cat(paste('\u2705', col, 'is empty \n'))
    } else {
      non_empty_rows<- which(!is.na(values) & values != "")
      non_empty_vals<- values[non_empty_rows]
      cat(paste('\u274C',col,'is NOT empty\n'))
      for( i in seq_along(non_empty_rows)){
        cat(paste('  Row',non_empty_rows[i],'>>',non_empty_vals[i], '\n'))
      }
      
    }
  }
}  
  

```{r}
# Q3/ Is sex coded correctly ? 


sex_var<- c('F','M','NA') 



for (df_name in names(df_lists)) {
  
  df <- df_lists[[df_name]]
  
  # Check if Sex column exists
  if ("Sex" %in% colnames(df)) {
    
    values <- df$Sex
    
    # Find invalid values
    invalid_values <- values[
      !is.na(values) &
        !(values %in% sex_var)
    ]
    
    if (length(invalid_values) > 0) {
      
      print(
        paste(
          df_name,
          "contains invalid Sex values:"
        )
      )
      
      print(unique(invalid_values))
      
    } else {
      
      print(
        paste(df_name, '\u2705',"- Sex column is valid")
      )
    }
    
  } else {
    
    print(
      paste(df_name,'\u26A0' ,"- Sex column NOT FOUND")
    )
  }
}


```

```{r}

# Q4/ Is ethnicity coded correctly?

ethnicity<-  c('A', 'AA', 'C', 'AS', 'H', 'O', 'NA') 
  
for (df_name in names(df_lists)) {
  df <- df_lists[[df_name]]
  
  if (!"Ethnicity" %in% colnames(df)) {
    cat(df_name, "- Ethnicity column NOT FOUND\n")
    next
  }
  
  invalid <- unique(df$Ethnicity[!is.na(df$Ethnicity) & !df$Ethnicity %in% ethnicity])   #Take invalid values
  
  if (length(invalid) == 0) {
    cat(df_name, "- ✅ All ethnicity values are valid\n")
  } else {
    cat(df_name, "- ❌ Invalid values:", paste(invalid, collapse = ", "), "\n")
  }
}



```



> The following function is desgined to take the varaible name, and its range to check each value in that varibale within the requried range or not. I used this function in Q5, Q6, Q8, Q9

```{r}

# Q5/ Are measurements date plausible? 

check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values variable
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'Day',       1,    31)
check_range(df_lists, 'Year',     1900,   2025)
check_range(df_lists, 'Month',     1,  12)

```
# Q6/ Are latitude, longitude and elevation plausible?? 

```{r}
#Q6/ Are latitude, longitude and elevation plausible?? 
check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values 
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'Latitude',       -90,    90)
check_range(df_lists, 'Longitude',     -180,   180)
check_range(df_lists, 'Elevation',     -430,  8850)



```

```{r}
#Q7/Is Country ISO valid?


ISO<- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Country ISO')%>%
  filter(if_any(everything(), ~ !is.na(.)))
  
ISO_country<- ISO$`Alpha-3 code`


for (df_name in names(df_lists)) {
  
  df     <- df_lists[[df_name]]
  values <- df$`Country ISO`
  
  if (any(values %in% ISO_country, na.rm = TRUE)) {
    cat("\u2705", df_name, ": Valid ISO country\n")
  } else {
    cat("\u274C", df_name, ": Invalid ISO country\n")
  }
}
```


```{r}


# Q8/Are age, weight and height within plausible range?
check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values 
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'Age',       0,    130)
check_range(df_lists, 'Dosing body weight',     0,   400)
check_range(df_lists, 'Height',     30,  250)

```


```{r}



# Q9/ Are ko, kd, No and Nd within plausible ranges?
check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values 
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'ko',       0.01,    2)
check_range(df_lists, 'kd',     0.01,   2)
check_range(df_lists, 'No',     50,  6000)
check_range(df_lists, 'Nd',     50,  6000)


```

```{r}
#Q10/ Is reproductive status coded correctly? 


cat('rep status female','\n___________________________\n')

female_status <- c("PP", "RANR", "P", "P p1", "P p2", "P p3", "L", "PM", "NA")
for (df_name in names(df_lists)) {
  df <- df_lists[[df_name]]
  
  if (!"rep status female" %in% colnames(df)) {
    cat(df_name, "- rep status female column NOT FOUND\n")
    next
  }
  
  invalid <- unique(df$`rep status female`[!is.na(df$`rep status female`) & !df$`rep status female` %in% female_status])   #Take invalid values
  
  if (length(invalid) == 0) {
    cat(df_name, "- ✅ All rep status female values are valid\n")
  } else {
    cat(df_name, "- ❌ Invalid values:", paste(invalid, collapse = ", "), "\n")
  }
}


male_status<-c("PP", "R", "NA")

cat('rep status male','\n___________________________\n')

for (df_name in names(df_lists)) {
  df <- df_lists[[df_name]]
  
  if (!"rep status male" %in% colnames(df)) {
    cat(df_name, "- rep status male column NOT FOUND\n")
    next
  }
  
  invalid <- unique(df$`rep status male`[!is.na(df$`rep status male`) & !df$`rep status male` %in% male_status])   #Take invalid values
  
  if (length(invalid) == 0) {
    cat(df_name, "- ✅ All rep status male values are valid\n")
  } else {
    cat(df_name, "- ❌ Invalid values:", paste(invalid, collapse = ", "), "\n")
  }
}


```

