---
title: "Quality control Pipline for the DLW Dataset"
output: 
  html_notebook:
    toc: true
    toc_float: true
    number_sections: true
---
## Readme for Non R users 
I will explain to you how to deal with this pipeline to do the quality control step for the DLW datasets.

The logic behind the pipeline can be divided into two steps: 1) load the data (Excel sheet), and 2) run the codes. Important: Before you load the file, make sure that the column names are EXACTLY the same as in the template because R is case-sensitive.

## Readme

This R Notebook provides a quality control (QC) pipeline for checking multiple Excel sheets from a merged DLW database submission file.

The pipeline loads 14 Excel sheets from one merged Excel workbook, stores them in a list of data frames, and runs a series of automated checks to identify missing values, invalid codes, incorrect formats, implausible values, and columns that should be empty.

The output is printed directly in the R console / R Notebook using clear status messages:

- ✅ Passed / valid / empty / within range
- ⚠️ Warning / missing values / not empty
- ❌ Failed / invalid / column not found / out of range

---

## Purpose of the Pipeline

The main purpose of this pipeline is to check whether DLW dataset sheets are ready for further review or submission.

It checks:

1. Whether all mandatory variables are filled in.
2. Whether autocalculated columns are empty.
3. Whether sex is coded correctly.
4. Whether ethnicity is coded correctly.
5. Whether measurement dates are plausible.
6. Whether latitude, longitude, and elevation values are plausible.
7. Whether Country ISO codes are valid.
8. Whether age, height, and body weight are within plausible ranges.
9. Whether ko, kd, No, and Nd values are within plausible DLW ranges.
10. Whether reproductive status variables are coded correctly.

---

## Important Requirements

This pipeline only works correctly if the column names in R match the column names in the Excel sheets *exactly*.

Before running the pipeline, make sure that:

1. All 14 Excel sheets use consistent column names.
2. The variables defined in R, such as `req_var` and `empty_col`, match the Excel column names exactly.
3. There are no extra spaces, spelling differences, or capitalization differences in the column names.

For example, these would be treated as different column names by R:

"Study ID"
"StudyID"
"study ID"
"Study ID "

--- 

#Workflow Start form here(non-R users):
This's the most important step for you because after that, you will only need to run the blocks and see the R results :).

1- First you have to set the working directory, this's important to tell R where are your files
2- Load the file
Here you only need to copy the file path of the file, and laod it into the R environment (Top right).
3- Add the file name to the df_lists.
for example if you file name is Pfrimer, add the name inside the df_lists. 
4- Run the block form the green right-arrow
```{r}

setwd("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data")


# Step 1: load the files 

#install.packages(c("tidyverse", "readxl"))

library(readxl) 
library(dplyr)


Parthenon <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Parthenon')%>%
  filter(if_any(everything(), ~ !is.na(.)))
Parthenon<- Parthenon[1:58,]


Kupad_Shivakumar <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Kupad_Shivakumar')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Kupad_Shivakumar<- Kupad_Shivakumar[1:89,]

Bangladesh <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Bangladesh')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Bangladesh<- Bangladesh[1:29,]


Bandara <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '196_Bandara')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Pfrimer <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '195_Pfrimer')%>%
  filter(if_any(everything(), ~ !is.na(.)))


Davy <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '194_Davy')%>%
  filter(if_any(everything(), ~ !is.na(.)))
Davy<- Davy[1:75,]


Giacomo_Fassini <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '192_Giacomo_Fassini')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Goldman <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '191_Goldman')%>%
  filter(if_any(everything(), ~ !is.na(.)))



weaver <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '190_Weaver(not_completed)')%>%
  filter(if_any(everything(), ~ !is.na(.)))



Desneves <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '193_Desneves')%>%
  filter(if_any(everything(), ~ !is.na(.)))


McGrosky  <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '189_McGrosky')%>%
  filter(if_any(everything(), ~ !is.na(.)))


Holsgrove_West<- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '188_Holsgrove-West')%>%
  filter(if_any(everything(), ~ !is.na(.)))

Holsgrove_West<- Holsgrove_West[1:25,]

silver<- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '187_Silver')%>%
  filter(if_any(everything(), ~ !is.na(.)))

silver<- silver[1:118,]

Pongcharoen <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = '186_Pongcharoen')%>%
  filter(if_any(everything(), ~ !is.na(.)))
Pongcharoen <- Pongcharoen[1:497,]


coker_RH <- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet= 'TEE+FILE+COKER+RH')%>%
  filter(if_any(everything(), ~ !is.na(.)))


# Step 2: merge all the lists together 

df_lists <- list(
  Bangladesh = Bangladesh,
  Bandara = Bandara,
  Davy = Davy,
  Desneves = Desneves,
  Giacomo_Fassini = Giacomo_Fassini,
  Goldman = Goldman,
  Holsgrove_West = Holsgrove_West,
  Kupad_Shivakumar = Kupad_Shivakumar,
  McGrosky = McGrosky,
  Parthenon = Parthenon,
  Pfrimer = Pfrimer,
  Pongcharoen = Pongcharoen,
  silver = silver,
  weaver = weaver,
  coker_RH = coker_RH)


```

>You can ignore the warning message.

> Clarification for the warning messages: Column AZ (column 52) was mostly empty at the top, so R guessed it was logical (TRUE/FALSE). Then at row 2716, it found the text "mass spec" — which R couldn't store as logical, so it likely converted it to NA.

# Q1/ Are all the mandatory variable filled in ? 

1- Here the logical steps are; 1- Create a variable that contains the names of the mandatory variables.
2- Go through each dataset, and if the req_var are present in one of the datset, check whether all values are filled in or not. 

> for Non-R USERS

1- You don't need to modify anything here, just run the block from the green right-arrow and see the results. 
```{r}

#Question 1 ---- 

# Q1/ Are all the mandatory variable filled in ? 

req_var <- c(
  "Study ID",
  "Individual ID",
  "Age",
  "Sex",
  "Dosing body weight",
  "ko",
  "kd",
  "No",
  "Nd",
  "rCO2",
  "rCO2 method",
  "TEE")

# First this assume that all the req_var is present in your df_lists




  for (df_name in names(df_lists)) {                      #Loop over the 14 dataset
    
    current_df <- df_lists[[df_name]]                     
    cat("\n📋 Checking:", df_name, "\n")
    
    for (var in req_var) {                               #2- loop over the req_var and check if the selected data has columnm in req_var
      
      if (var %in% colnames(current_df)) {                    #3- if req_var in one of the df 
        
        values <- current_df[[var]]           # ✅ keep all values including NAs
        
        if (all(!is.na(values))) {            # ✅ check if ALL values are filled
          cat("  \u2705 Fully filled", var, "- Fully filled \n")
        } else {
          n_missing <- sum(is.na(values))     # ✅ count the NAs
          cat("  ⚠️ ", var, "-", n_missing, "missing values \n")
        }
        
      }  else {
        cat("  ❌", var, "- Column not found \n")
      }
    }
  }




```

> Note: Pongcharoen Study ID is completely missing.


> Heatmap to see the % of the completeness

```{r}
library(tidyverse)

# Build a results dataframe from your loop
results <- map_dfr(names(df_lists), function(df_name) {
  current_df <- df_lists[[df_name]]
  map_dfr(req_var, function(var) {
    if (!var %in% colnames(current_df)) {                    #If the variable is missing, put absent in status
      tibble(dataset = df_name, variable = var, status = "absent", pct_complete = NA)
    } else {
      values <- current_df[[var]]                      #If not missing take the values
      n_total   <- length(values)                      # count the values 
      n_missing <- sum(is.na(values))                   #Count the missing values 
      tibble(                                            #create results dataframe 
        dataset     = df_name,
        variable    = var,
        status      = ifelse(n_missing == 0, "ok", "warn"),            #If the missing zero put ok, else warn 
        pct_complete = round((1 - n_missing / n_total) * 100)           #count the %
      )
    }
  })
})

# Plot the heatmap
results %>%                              #Take the df results and make label column to 
  mutate(label = case_when(
    status == "absent" ~ "N/A",
    pct_complete == 100 ~ "100%",
    TRUE ~ paste0(pct_complete, "%")     #for all cases write the value pct_com and put %
  )) %>%
  ggplot(aes(x = variable, y = dataset, fill = pct_complete)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient2(low = "#E24B4A", mid = "#FAC775", high = "#1D9E75",
                       midpoint = 50, na.value = "#D3D1C7", limits = c(0, 100),
                       name = "% complete") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = NULL, title = "QC completeness across datasets")
```




# Q2/ Are all autocalculated columns empty? 

 - Here, the logical steps are:
1- Create a variable that contains the column names that must be empty.
2- Go through each dataset.
3- If the columns in empty_col match the column names in the dataset, check whether they are empty or not.
3- If all the columns are empty in the dataset, print the dataset name and state that they are empty. If not, extract the values and show their location in the dataset.


```{r}
# Q2/ Are all autocalculated columns empty? 

empty_col<- c("Database measurement ID","Database individual ID","BMI","Average weight",
              "Weight change","ko/kd","Nd/No","pred DSR","autocompleted CO2 production","TBW","Autocompleted TEE Speakman et al 2021","Water turnover","Hydration",
              "FFM (kg)","FFM(0.732)","FM (kg)","Adiposity","Analysis lab anon")


for( df_name in names(df_lists)){
  df<- df_lists[[df_name]]
  matched_cols<- intersect(colnames(df),empty_col) #find the empty_cols in the datasets
  
  if(length(matched_cols) == 0) next             #skip if the matched _cols == zero 
  cat('\n',df_name,'\n')
   cat('---------------------','\n')
  for(col in matched_cols){
    values<-df[[col]]                           #extract the column values
    
    if(all(is.na(values) | values == "")) {
      
      cat(paste('\u2705', col, 'is empty \n'))
    } else {
      non_empty_rows<- which(!is.na(values) & values != "")
      non_empty_vals<- values[non_empty_rows]
      cat(paste('\u274C',col,'is NOT empty\n'))
      for( i in seq_along(non_empty_rows)){
        cat(paste('  Row',non_empty_rows[i],'>>',non_empty_vals[i], '\n'))
      }
      
    }
  }
}  
  


```

# Q3/ Is sex coded correctly ? 
- Here, the logical steps are:
1- Create a variable that contains the accepted sex values: F, M, and NA.
2- Go to the Sex column in each dataset.
3- Check whether all values match one of the accepted sex values.

```{r}
# Q3/ Is sex coded correctly ? 


sex_var<- c('F','M','NA') 



for (df_name in names(df_lists)) {
  
  df <- df_lists[[df_name]]
  
  # Check if Sex column exists
  if ("Sex" %in% colnames(df)) {
    
    values <- df$Sex
    
    # Find invalid values
    invalid_values <- values[
      !is.na(values) &
        !(values %in% sex_var)
    ]
    
    if (length(invalid_values) > 0) {
      
      print(
        paste(
          df_name,
          "contains invalid Sex values:"
        )
      )
      
      print(unique(invalid_values))
      
    } else {
      
      print(
        paste(df_name, '\u2705',"- Sex column is valid")
      )
    }
    
  } else {
    
    print(
      paste(df_name,'\u26A0' ,"- Sex column NOT FOUND")
    )
  }
}


```



# Q4/ Is ethnicity coded correctly?
- Here, the logical steps are:
1- Create a variable that contains the accepted ethnicity codes.
2- Go to the Ethnicity column in each dataset.
3- Check whether all values match one of the accepted ethnicity codes.
4- If there are invalid values, print the dataset name and show the invalid values.

```{r}

# Q4/ Is ethnicity coded correctly?

ethnicity<-  c('A', 'AA', 'C', 'AS', 'H', 'O', 'NA') 
  

#This for loop would take all the column values and compare it to the ethnicity, if something is not valid, it will print invalid values 
for (df_name in names(df_lists)) {
  df <- df_lists[[df_name]]
  
  if (!"Ethnicity" %in% colnames(df)) {
    cat(df_name, "- Ethnicity column NOT FOUND\n")
    next
  }
  
  invalid <- unique(df$Ethnicity[!is.na(df$Ethnicity) & !df$Ethnicity %in% ethnicity])   #Take invalid values
  
  if (length(invalid) == 0) {
    cat(df_name, "- ✅ All ethnicity values are valid\n")
  } else {
    cat(df_name, "- ❌ Invalid values:", paste(invalid, collapse = ", "), "\n")
  }
}



```



> The following function is desgined to take the varaible name, and its range to check each value in that varibale within the requried range or not. I used this function in Q5, Q6, Q8, Q9


# Q5/ Are measurements date plausible? 
- Here, the logical steps are:
1- Create a function that take input (datasets,colname(variables), ranges, label).
2- define each input in the function.
3- Add your input to the function and see the output.

```{r}

# Q5/ Are measurements date plausible? 

check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values variable
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'Day',       1,    31)
check_range(df_lists, 'Year',     1900,   2025)
check_range(df_lists, 'Month',     1,  12)

```


> Explaination: R printed NA as values because in the main dataset, NA is written as text rather than being an empty cell.

# Q6/ Are latitude, longitude and elevation plausible?? 

```{r}
#Q6/ Are latitude, longitude and elevation plausible?? 
check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values 
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'Latitude',       -90,    90)
check_range(df_lists, 'Longitude',     -180,   180)
check_range(df_lists, 'Elevation',     -430,  8850)



```

# Q7/ Is Country ISO valid?

- Here, the logical steps are:
1- Create a variable that contains the ISO code.
2- Go to the Country ISO in each dataset. 
3- If onE of the ISO code is present in the dataset.
4- Print Valid ISO country.
5- If not, print Invalid ISO country.

```{r}
#Q7/Is Country ISO valid?


ISO<- read_excel("C:/Users/MOHAMEDAH/OneDrive - IAEA/Ahmed/Nutrition_health/DLW_Alexia/draft_data/template.xlsx", sheet = 'Country ISO')%>%
  filter(if_any(everything(), ~ !is.na(.)))
  
ISO_country<- ISO$`Alpha-3 code`


for (df_name in names(df_lists)) {
  
  df     <- df_lists[[df_name]]
  values <- df$`Country ISO`
  
  if (any(values %in% ISO_country, na.rm = TRUE)) {
    cat("\u2705", df_name, ": Valid ISO country\n")
  } else {
    cat("\u274C", df_name, ": Invalid ISO country\n")
  }
}
```

# Q8/Are age, weight and height within plausible range?


```{r}


# Q8/Are age, weight and height within plausible range?
check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values 
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'Age',       0,    130)
check_range(df_lists, 'Dosing body weight',     0,   400)
check_range(df_lists, 'Height',     30,  250)

```



# Q9/ Are ko, kd, No and Nd within plausible ranges?

```{r}
# Q9 Are ko, kd, No and Nd within plausible ranges?


# Q8/Are age, weight and height within plausible range?
check_range <- function(df_lists, col_name, min_val, max_val, label = col_name) {
  cat('\n', label, '\n--------------------\n')    # 1)print the label
  for (df_name in names(df_lists)) {               # 2) loop over the df_lists
    df <- df_lists[[df_name]]                       # 3) take the dataset 
    if (!col_name %in% names(df)) next                # 4) If col_name not in the colnames of the data, skip 
    values <- suppressWarnings(as.numeric(df[[col_name]]))   # 5) Take the column values and put it into values 
    if (all(is.na(values))) {                                  
      cat('\u26A0',df_name, ':', label, 'column is empty\n')
    } else if (all(values >= min_val & values <= max_val, na.rm = TRUE)) {     # 6) Here check the values range
      cat('\u2705',df_name, ': Within the', label, 'range\n')
    } else {
      invalid <- values[!is.na(values) & (values < min_val | values > max_val)]                    #7) if not within the range take the invalid values
      cat('\u274C',df_name, ': Out of range →', paste(invalid, collapse = ", "), '\n')             #8) Tell me where are these values
    }
  }
}

check_range(df_lists, 'ko',       0.01,    2)
check_range(df_lists, 'kd',     0.01,   2)
check_range(df_lists, 'No',     50,  6000)
check_range(df_lists, 'Nd',     50,  6000)


```
> Note:In Bandara and Pongcharoen datasets, NOT ALL the values are within the required range)



# Q10/ Is reproductive status coded correctly? 

```{r}
#Q10/ Is reproductive status coded correctly? 


cat('rep status female','\n___________________________\n')

female_status <- c("PP", "RANR", "P", "P p1", "P p2", "P p3", "L", "PM", "NA")
for (df_name in names(df_lists)) {
  df <- df_lists[[df_name]]
  
  if (!"rep status female" %in% colnames(df)) {
    cat(df_name, "- rep status female column NOT FOUND\n")
    next
  }
  
  invalid <- unique(df$`rep status female`[!is.na(df$`rep status female`) & !df$`rep status female` %in% female_status])   #Take invalid values
  
  if (length(invalid) == 0) {
    cat(df_name, "- ✅ All rep status female values are valid\n")
  } else {
    cat(df_name, "- ❌ Invalid values:", paste(invalid, collapse = ", "), "\n")
  }
}



male_status<-c("PP", "R", "NA")

cat('rep status male','\n___________________________\n')

for (df_name in names(df_lists)) {
  df <- df_lists[[df_name]]
  
  if (!"rep status male" %in% colnames(df)) {
    cat(df_name, "- rep status male column NOT FOUND\n")
    next
  }
  
  invalid <- unique(df$`rep status male`[!is.na(df$`rep status male`) & !df$`rep status male` %in% male_status])   #Take invalid values
  
  if (length(invalid) == 0) {
    cat(df_name, "- ✅ All rep status male values are valid\n")
  } else {
    cat(df_name, "- ❌ Invalid values:", paste(invalid, collapse = ", "), "\n")
  }
}


```

